New and simplified manual controls for projection and slice tours, with application to exploring classification boundaries in high dimensions

### 2023-02-08 review/response.Rmd

This paper describes a method for interacting with multidimensional data using slice tours in Mathematica (as well as a reduced-feature implementation in R).

Major Issues

- Reproducibility:
    - Code should be reproducible and not depend on user paths.
    - Data files should be uploaded along with code files (they don't seem to be part of supplemental material?)
    - Not having the ability to actually run the code (because the data wasn't included) makes it somewhat difficult to assess the software side of this paper.

\textcolor{blue}{Done. Previously we provided the public github repo URL so that all files to reproduce could be downloaded. Here we have included all the files in the supplementary zip.}

### 2023-02-08 paper/paper.Rmd
removed an empty line
note on standardiisation
Changes working on review

To better explore the space we combine the manual controls for the projection with manual controls for slicing. A slice is a section of the data that is defined by a projection, a center point that is anchoring it in the high-dimensional space, and the slice thickness $h$ [@slicetour]. A data point is inside the slice if its orthogonal distance from the projection plane (passing through the center point) is below the thickness $h$. This orthogonal distance is computed in terms of the component that is normal on the projection plane. For $\mathbf{x}_i$ a $p$ dimensional data point and $\mathbf{c}$ the center point (in the same $p$ dimensional space) we compute the orthogonal distance as
\begin{equation}
v_i^2 = ||\mathbf{x}_i' - \mathbf{c}'||^2 = \mathbf{x}_i'^2 + \mathbf{c}'^2 - 2 \mathbf{x}_i'\cdot\mathbf{c}' ,
\label{eq:slice}
\end{equation}
with $\mathbf{c}' = \mathbf{c} - (\mathbf{c}\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{c}\cdot \mathbf{a}_2 )\mathbf{a}_2$, $\mathbf{x}_i' = \mathbf{x}_i - (\mathbf{x}_i\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{x}_i\cdot \mathbf{a}_2) \mathbf{a}_2$ and $\mathbf{a}_k, k=1,2 (=d)$ denoting the columns of the projection matrix, $\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)$.

@@

The source material and animated gifs for this paper are available at [https://github.com/uschiLaa/mmtour](https://github.com/uschiLaa/mmtour).

The supplementary materials include:

- The Mathematica source code defining the new functions in mmtour.wl.
- Three Mathematica notebooks with the code used in the application (corresponding to the three subsections).
- Appendix with additional details about manual controls, and Mathematica functions.
- Data files for the examples.
- R code to reproduce the example data, Figures 1 and 2 in the paper and Figures 1 and 2 in the appendix.
- Animations illustrating the manual tour and slicing, matching the static figures in the paper. These are also available at https://vimeo.com/747585410 and https://vimeo.com/747590472.

>>>>

To better explore the space we combine the manual controls for the projection with manual controls for slicing. A slice is a section of the data that is defined by a projection, a center point that is anchoring it in the high-dimensional space, and the slice thickness $h$ [@slicetour]. A data point is inside the slice if its orthogonal distance from the projection plane (passing through the center point) is below the thickness $h$. This orthogonal distance is computed in terms of the component that is normal on the projection plane. For $\mathbf{x}_i$ a $p$ dimensional data point and $\mathbf{c}$ the center point (in the same $p$ dimensional space) we compute the orthogonal distance as
\begin{equation}
v_i^2 = ||\mathbf{x}_i' - \mathbf{c}'||^2 = \mathbf{x}_i'^2 + \mathbf{c}'^2 - 2 \mathbf{x}_i'\cdot\mathbf{c}' ,
\label{eq:slice}
\end{equation}
with $\mathbf{c}' = \mathbf{c} - (\mathbf{c}\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{c}\cdot \mathbf{a}_2 )\mathbf{a}_2$, $\mathbf{x}_i' = \mathbf{x}_i - (\mathbf{x}_i\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{x}_i\cdot \mathbf{a}_2) \mathbf{a}_2$ and $\mathbf{a}_k, k=1,2 (=d)$ denoting the columns of the projection matrix, $\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)$. Although not required, manual control of slices works more understandably if the data is first standardized.

@@

The source material and animated gifs for this paper are available at [https://github.com/uschiLaa/mmtour](https://github.com/uschiLaa/mmtour). The supplementary materials include:

- The Mathematica source code defining the new functions in mmtour.wl.
- Two Mathematica notebooks with the code used in the application (corresponding to the three subsections).
- Appendix with additional details about manual controls, and Mathematica functions.
- Data files for the examples.
- R code to reproduce the example data, Figures 1 and 2 in the paper and Figures 1 and 2 in the appendix.
- Animations illustrating the manual tour and slicing are also available at https://vimeo.com/747585410 and https://vimeo.com/747590472.

### 2023-02-08 paper/paper.tex
removed an empty line
note on standardiisation
Changes working on review

To better explore the space we combine the manual controls for the
projection with manual controls for slicing. A slice is a section of the
data that is defined by a projection, a center point that is anchoring
it in the high-dimensional space, and the slice thickness \(h\)
\citep{slicetour}. A data point is inside the slice if its orthogonal
distance from the projection plane (passing through the center point) is
below the thickness \(h\). This orthogonal distance is computed in terms
of the component that is normal on the projection plane. For
\(\mathbf{x}_i\) a \(p\) dimensional data point and \(\mathbf{c}\) the
center point (in the same \(p\) dimensional space) we compute the
orthogonal distance as \begin{equation}
v_i^2 = ||\mathbf{x}_i' - \mathbf{c}'||^2 = \mathbf{x}_i'^2 + \mathbf{c}'^2 - 2 \mathbf{x}_i'\cdot\mathbf{c}' ,
\label{eq:slice}
\end{equation} with
\(\mathbf{c}' = \mathbf{c} - (\mathbf{c}\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{c}\cdot \mathbf{a}_2 )\mathbf{a}_2\),
\(\mathbf{x}_i' = \mathbf{x}_i - (\mathbf{x}_i\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{x}_i\cdot \mathbf{a}_2) \mathbf{a}_2\)
and \(\mathbf{a}_k, k=1,2 (=d)\) denoting the columns of the projection
matrix, \(\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)\).

@@

The source material and animated gifs for this paper are available at
\url{https://github.com/uschiLaa/mmtour}.

The supplementary materials include:

\begin{itemize}
\tightlist
\item
  The Mathematica source code defining the new functions in mmtour.wl.
\item
  Three Mathematica notebooks with the code used in the application
  (corresponding to the three subsections).
\item
  Appendix with additional details about manual controls, and
  Mathematica functions.
\item
  Data files for the examples.
\item
  R code to reproduce the example data, Figures 1 and 2 in the paper and
  Figures 1 and 2 in the appendix.
\item
  Animations illustrating the manual tour and slicing, matching the
  static figures in the paper. These are also available at
  \url{https://vimeo.com/747585410} and
  \url{https://vimeo.com/747590472}.
\end{itemize}

>>>>

To better explore the space we combine the manual controls for the
projection with manual controls for slicing. A slice is a section of the
data that is defined by a projection, a center point that is anchoring
it in the high-dimensional space, and the slice thickness \(h\)
\citep{slicetour}. A data point is inside the slice if its orthogonal
distance from the projection plane (passing through the center point) is
below the thickness \(h\). This orthogonal distance is computed in terms
of the component that is normal on the projection plane. For
\(\mathbf{x}_i\) a \(p\) dimensional data point and \(\mathbf{c}\) the
center point (in the same \(p\) dimensional space) we compute the
orthogonal distance as \begin{equation}
v_i^2 = ||\mathbf{x}_i' - \mathbf{c}'||^2 = \mathbf{x}_i'^2 + \mathbf{c}'^2 - 2 \mathbf{x}_i'\cdot\mathbf{c}' ,
\label{eq:slice}
\end{equation} with
\(\mathbf{c}' = \mathbf{c} - (\mathbf{c}\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{c}\cdot \mathbf{a}_2 )\mathbf{a}_2\),
\(\mathbf{x}_i' = \mathbf{x}_i - (\mathbf{x}_i\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{x}_i\cdot \mathbf{a}_2) \mathbf{a}_2\)
and \(\mathbf{a}_k, k=1,2 (=d)\) denoting the columns of the projection
matrix, \(\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)\). Although not
required, manual control of slices works more understandably if the data
is first standardized.

@@

The source material and animated gifs for this paper are available at
\url{https://github.com/uschiLaa/mmtour}. The supplementary materials
include:

\begin{itemize}
\tightlist
\item
  The Mathematica source code defining the new functions in mmtour.wl.
\item
  Two Mathematica notebooks with the code used in the application
  (corresponding to the three subsections).
\item
  Appendix with additional details about manual controls, and
  Mathematica functions.
\item
  Data files for the examples.
\item
  R code to reproduce the example data, Figures 1 and 2 in the paper and
  Figures 1 and 2 in the appendix.
\item
  Animations illustrating the manual tour and slicing are also available
  at \url{https://vimeo.com/747585410} and
  \url{https://vimeo.com/747590472}.
\end{itemize}
